Bioverse: The Habitable Zone Inner Edge Discontinuity as an Imprint of Runaway Greenhouse Climates on Exoplanet Demographics

### 2023-05-11 src/tex/ms.tex
add minor edits

Nominally, the above models assume a \textit{dry melt} without dissolved volatiles.
Here, however, we consider a \textit{wet melt} magma ocean and take into account a radius decrease from retention of water in the melt~\citep{Dorn2021}.
The impact of the water distribution between melt and atmosphere on the change of the transit radius depends on the planet's mass and water content and is generally small compared to the radius inflation from the steam atmosphere.
%We follow \citet{Dorn2021} and use their computed radius deviations between a wet magma ocean and a solid mantle for a tropopause pressure~$P_\mathrm{iso}=\SI{0.1}{\bar}$.
We computed radius deviations between a wet magma ocean and a solid mantle following \citet{Dorn2021} assuming a tropopause pressure~$P_\mathrm{iso}=\SI{0.1}{\bar}$.
We then added the (in almost all cases negative) radius deviations to the inflated planet radii computed for the dry melt case.

@@

\subsubsection{Water mass fraction}
The predominant mass fractions of water, which sensitively controls the atmospheric state of a rocky exoplanet, is poorly constrained.
Inferred water contents in the literature range from upper limits on the order $10^{-5}$ to ``water worlds'' with tens of percent mass fraction~\citep[e.g.,][]{Rogers2010,Unterborn2018,Mousis2020,Agol2021,Luque2022}, all of which are within the realm of theoretical predictions~\citep{Mulders2015b,Sato2016,Jin2018,Lichtenberg2019,Bitsch2019b,Venturini2020,Emsenhuber2021b,Schlecker2021,2022ApJ...938L...3L,2022ApJ...939L..19I}.
Our nominal case assumes a bulk water mass fraction of $x_\mathrm{H_2O} = \var{wrr}$.
This can be considered a conservative choice that is unlikely to introduce a systematic overestimation of the habitable zone inner edge discontinuity. %, and the hypothesis tests are only marginally sensitive to $x_\mathrm{H_2O}$ as long as it is $\gtrsim 10^{-4}$ (see Sect.~\ref{sec:res_testability}).
Cases of pure rocky composition and very low volatile contents can be considered absorbed by the dilution factor $f_\mathrm{rgh}$.
Assuming a distribution of water mass fractions instead of a fixed value would thus not significantly change our results.

>>>>

Nominally, the above models assume a \textit{dry melt} without dissolved volatiles.
Here, however, we consider a \textit{wet melt} magma ocean and take into account a radius decrease from retention of water in the melt~\citep{Dorn2021}.
The impact of the water distribution between melt and atmosphere on the change of the transit radius depends on the planet's mass and water content and is generally small compared to the radius inflation from the steam atmosphere.
%We follow \citet{Dorn2021} and use their computed radius deviations between a wet magma ocean and a solid mantle for a tropopause pressure~$P_\mathrm{iso}=\SI{0.1}{\bar}$.
We computed radius deviations between a wet magma ocean and a solid mantle following \citet{Dorn2021} and assuming a tropopause pressure~$P_\mathrm{iso}=\SI{0.1}{\bar}$.
We then added the (in almost all cases negative) radius deviations to the inflated planet radii computed for the dry melt case.

@@

\subsubsection{Bulk water mass fraction}
The predominant mass fractions of water, which sensitively controls the atmospheric state of a rocky exoplanet, is poorly constrained.
Inferred water contents in the literature range from upper limits on the order $10^{-5}$ to ``water worlds'' with tens of percent mass fraction~\citep[e.g.,][]{Rogers2010,Unterborn2018,Mousis2020,Agol2021,Luque2022}, all of which are within the realm of theoretical predictions~\citep{Mulders2015b,Sato2016,Jin2018,Lichtenberg2019,Bitsch2019b,Venturini2020,Emsenhuber2021b,Schlecker2021,2022ApJ...938L...3L,2022ApJ...939L..19I}.
Our nominal case assumes a bulk water mass fraction of $x_\mathrm{H_2O} = \var{wrr}$.
This can be considered a conservative choice that is unlikely to introduce a systematic overestimation of the habitable zone inner edge discontinuity. %, and the hypothesis tests are only marginally sensitive to $x_\mathrm{H_2O}$ as long as it is $\gtrsim 10^{-4}$ (see Sect.~\ref{sec:res_testability}).
Cases of pure rocky composition and very low volatile contents can be considered absorbed by the dilution factor $f_\mathrm{rgh}$.
Assuming a distribution of water mass fractions instead of a fixed value would thus not significantly change our results.
